Supplementary Materials

### 2022-12-08 manuscript/supplementary.tex
more notes

Our underlying data for all likelihood and pairwise diversity estimates is the
allele count matrix $\mathbf{C}$ with dimensions $l \times 2$, where $l$ is the
chromosome length. This is converted to a pairwise summary matrix with
identical dimensions, $\mathbf{P}$. The first column of $\mathbf{P}$ is the
number of pairwise comparisons between chromosomes that are identical, and the
second column is the number that are different. Both of these columns are
combinatoric summaries of the allele counts matrix. Let $[c_1, c_2]$ be a row
of $\mathbf{C}$ for basepair $i$, and $n = c_1 + c_2$. Then, the (1) total
number of pairwise combinations of chromosomes $n_T$, (2) the number of
pairwise with identical alleles $n_S$, and (3) the number of pairwise
combinations with differing alleles $n_D$ are respectively,

@@

Thus $\mathbf{P}_i = [n_S, n_D]$, and the pairwise diversity is calculated as

\begin{align}
  \bar{\pi}(i) = \frac{n_D}{n_T}
\end{align}

@@

\begin{align}
  \bar{\pi}(i) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
\end{align}

@@

>>>>

Our underlying data for all likelihood and pairwise diversity estimates is the
allele count matrix $\mathbf{C}$ with dimensions $L \times 2$, where $L$ is the
chromosome length. This is converted to a pairwise summary matrix with
identical dimensions, $\mathbf{P}$. The first column of $\mathbf{P}$ is the
number of pairwise comparisons between chromosomes that are identical, and the
second column is the number that are different. Both of these columns are
combinatoric summaries of the allele counts matrix. Let $[c_1, c_2]$ be a row
of $\mathbf{C}$ for basepair $l$, and $n = c_1 + c_2$. Then, the (1) total
number of pairwise combinations of chromosomes $n_T$, (2) the number of
pairwise with identical alleles $n_S$, and (3) the number of pairwise
combinations with differing alleles $n_D$ are respectively,

@@

Thus $\mathbf{P}_l = [n_S, n_D]$, and the pairwise diversity is calculated as

\begin{align}
  \bar{\pi}(l) = \frac{n_D}{n_T}
\end{align}

@@

\begin{align}
  \bar{\pi}(l) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
\end{align}

@@

This per-basepair diversity is then averaged over all callable bases, to give a
genome-wide or window estimate of diversity of accessible bases in the set
$\mathcal{A}$,

\begin{align*}
  \bar{\pi}(\mathcal{A}) = \frac{\sum_{l \in \mathcal{A}} \bar{\pi}(l)}{L_\mathcal{A}}.
\end{align*}

where $L_\mathcal{A} = |\mathcal{A}|$ is the number of accessible bases. Note
that both $L_\mathcal{A}$ and the sum are random variables in genomic data.
The denominator varies along the genome, based on the number of
reliably-genotyped and accessible bases in the reference genome. This is a
ratio estimator, and is thus biased, since by Jensen's inequality
$\E(\nicefrac{y}{x}) \ge \nicefrac{\E(y)}{\E(x)}$.

We explore this with block-bootstrap bias correction. TODO
